---
title: "package-walkthrough"
output: rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{package-walkthrough}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup}

library(habcontool)
library(tidyverse)
library(patchwork)
library(sf)

```

# Habitat Connectivity tool

This is an R package written for the ERAMMP and EVAST projects which is used to is used to determine spatial locations in which planting new trees/introducing new habitats would connect two or more habitat patches together. It relies heavily on the excellent _sf_ (https://r-spatial.github.io/sf/) and _terra_ (https://rspatial.github.io/terra/) packages. 

## Brief overview

In brief, this package buffers polygons from the _sf_ package, identifies areas where polygons overlap and returns a raster showing overlapping areas. This is to determine spatial locations in which introducing more of a given habitat would increase connectivity between existing habitat patches. This code can be used for any habitat, or anything really, so long as it is given in polygon format. The functions identify locations in which the addition of a habitat patch would connect two or more habitat patches. 

Here, we provide a brief overview of the main package functions. We do this using data supplied in the _sf_ package. These are states in the US but we will pretend they represent a mosaic of patches of woodland. 

```{r data-setup}

#> load data from sf package
sf_obj <- st_read(system.file("shape/nc.shp", package="sf"))

#> randomly sample rows
habs <- sf_obj[sample(1:nrow(sf_obj), 20),]
plot(st_geometry(habs))
habs

# check coordinates
st_crs(habs)

# let's transform to something a bit more normal
habs <- st_transform(habs, "EPSG:4087")

```


There are two main functions in the _habcontool_ package, `habitat_overlap` and `habitat_overlap_gridded`. For most scenarios, `habitat_overlap` will work fine. `habitat_overlap_gridded` should really only be used when there are many complex polygons at a fine spatial resolution. Identifying overlaps between many polygons with complex shapes (i.e. think many very oddly shaped polygons close together) is computationally challenging. Gridding up the area and looking for overlaps within each grid is a simple way of reducing the computational load - and this is what `habitat_overlap_gridded` does.

First, we'll explain the `habitat_overlap` function. This function takes an _sf_ dataframe of spatial objects, buffers them, finds overlaps and returns a _spatRast_ from the _terra_ package. 

### Parameters and what they do

* `spatial_object`: the spatial object that you want to use to investigate habitat connectivity with.

* `habitat_column_name`: the name of the column in your dataframe that specifies the habitat. This is useful if you have multiple habitats in your polygons and want to know which ones have been "connected" if `combine_touching_polys` and/or `combine_close_polys` are `TRUE`.

* `extent`: a vector (`xmin`, `ymin`, `xmax`, `ymax`) specifying a subset of your domain if you only want to run it on a small area.

* `buffer_distance`: the distance that you want to buffer each habitat patch by. This is the maximum distance at which you consider two habitat patches to be connected after the introduction of new habitats has occurred. 

* `connection_distance`: only used when `combine_touching_polys` and `combine_close_polys` are `TRUE`. The distance at which habitat patches are already considered to be the same patch. Useful in the case of highly mobile species that use habitat mosaics.

* `min_area`: the minimum area of habitat patches you want to consider for buffering.

* `combine_touching_polys`: whether you want to combine habitat patches that touch each other.

* `combine_close_polys`: whether you want to consider habitat patches that are within `connection_distance` of each other already connected.

* `plot_it`: should the function return plots.

* `resolution`: the resolution of the raster that you want returned. This should be carefully thought about and tested before running on a large area.


###### Find a better shapefile

### Running `habitat_overlap`

Here we'll look at the running of the `habitat_overlap` function. There are a few parts of the function call that are worth noting. First, we have set `buffer_distance = 100000`. This means that we are buffering each polygon by 100000m and trying to find areas where there is overlap between them. Introducing new habitat in these overlapping regions means that each of the original patches will have habitat within 100000m of each other. Second, we set `connection_distance = 500`. This distance is specifically for when `combine_touching_polys` and `combine_close_polys` are `TRUE`. This is the distance at which patches, before buffering, are already considered connected. Polygons within this distance of each other are considered as a single polygon for the identification of overlapping habitats. Finally, we set `resolution = c(1000, 1000)`. This is the resolution of the output raster (c(x coordinate, y coordinate)), and is on the scale of the spatial object we are buffering. I.e. the spatial object we are using here is on a WGS84 coordinate system, which is measured in metres. The output raster will therefore be in 10km increments. It is important to know what coordinate system you are working in - get more information using this website https://epsg.io/.

Ultimately, the values that these parameters take will be heavily influenced by the project questions. It is worth testing different values to understand how these will affect the results. 


```{r hab-overlap}

overs <- habitat_overlap(spatial_object = habs, 
                         habitat_column_name = NULL, 
                         extent = NULL, 
                         buffer_distance = 100000,
                         min_area = NULL, 
                         combine_touching_polys = TRUE, 
                         combine_close_polys = TRUE,
                         connection_distance = 500, 
                         plot_it = TRUE, 
                         resolution = c(10000,10000))

```


If `plot_it = TRUE`, four plots are plotted when running the function. These show the initial object, the objects after combining has occurred (irrespective of whether `combine_touching_polys` and `combine_close_polys` are `TRUE`), and the overlapping regions only. 

The output is a list of five objects from different steps in the processing. The fifth entry is the raster, at the resolution specified, which identifies overlapping regions. The value of cells specifies the number of habitat patches that would be connected if planting happened in this area.

It's worth playing around with the different parameters to see how it affects your results. As you'll see, the values you choose will have quite a large effect.For example, let's change the buffer to a smaller value, representing a situation where we are interested in connectivity over a smaller distance.

```{r small-buff}

overs <- habitat_overlap(spatial_object = habs, 
                         habitat_column_name = NULL, 
                         extent = NULL, 
                         buffer_distance = 25000,
                         min_area = NULL, 
                         combine_touching_polys = TRUE, 
                         combine_close_polys = TRUE,
                         connection_distance = 500, 
                         plot_it = TRUE, 
                         resolution = c(1000,1000))

```


### Running `habitat_overlap_gridded`

In some cases, we are interested in running connectivity across a large spatial scale (perhaps with many complex polygons). To do this, we can use the `habitat_overlap_gridded` function, which works by gridding the whole object up and working on smaller subsets at a time. For ease, we will use the spatial dataset above as an example. This will be slower that running the `habitat_overlap` function by itself. However, for bigger and more complex shapefiles, the `habitat_overlap_gridded` function will be quicker. 

Before running the function, we first need to create a spatial grid to work over the area of interest. 

buffer should be the same as the connection distance we are looking to find 

```{r create-grid}

grids <- create_grids(spatial_object = habs, grid_size = 80000, grid_buffer_size = 50000)

```


```{r}


### Need to try and combine this stuff into the habitat_overlap_gridded function


habitat_overlap_gridded(spatial_object = habs, 
                        SQL_query = NULL, 
                        habitat_column_name = NULL, 
                        buffer_distance = 25000, 
                        min_hab_area = NULL, 
                        combine_touching_polys = TRUE,
                        combine_close_polys = TRUE, 
                        connection_distance = 500,
                        plot_it = FALSE, 
                        resolution = c(1000,1000),
                        extent_large = grids[[2]],
                        extent_central = grids[[1]], 
                        save = FALSE, 
                        save_loc, 
                        save_name)


gridded_overlaps_out <- list()

for(i in 1:length(grids[[1]])){
  print(i)
  
  tryCatch(
    {
      gridded_overlaps_out[[i]] <- 
        suppressMessages(habitat_overlap_gridded(spatial_object = habs, 
                                                 SQL_query = NULL, 
                                                 habitat_column_name = NULL, 
                                                 buffer_distance = 25000, 
                                                 min_hab_area = NULL, 
                                                 combine_touching_polys = TRUE,
                                                 combine_close_polys = TRUE, 
                                                 connection_distance = 500,
                                                 plot_it = FALSE, 
                                                 resolution = c(1000,1000),
                                                 extent_large = grids[[2]],
                                                 extent_central = grids[[1]], 
                                                 save = FALSE, 
                                                 save_loc, 
                                                 save_name))
    },
    
    error = function(cond) warning(cond),
    finally = next()
    
  )
  
}

```

```{r}

allgrids <- do.call(rbind, gridded_overlaps_out)

combgrid <- allgrids %>% 
  summarise(geometry = st_union(geometry)) %>% 
  st_cast("POLYGON")

plot(st_geometry(allgrids))
plot((overs$habitat_connectivity_raster))
plot(st_geometry(combgrid))

```


TWO MORE SECTIONS = 

One on the habitat overlap gridded - write up new function to combine the gridded objects too

One on how to write up the submission file for lotus (specifically for Amy and Freya...)




